Unifying quasars with clumpy wind models

### 2015-08-27 paper_mnras.tex
version for sending to collabs

\maketitle
\begin{abstract} 
% Broad absorption lines (BALs) in the ultraviolet 
% are seen in $\sim20\%$ of quasi-stellar objects (QSOs). 
% Blue-shifted broad absorption lines (BALs) are the most direct evidence of 
% accretion disc `winds' in such systems; mass loaded outflows
% emanating from the disc that may be driven by line forces or
% magnetic processes. 
Various unification schemes for
quasars and luminous active galactic nuclei (AGN) have proposed
that the broad emission line region is roughly cospatial
with broad absorption line (BAL) gas and much of the phenomenology of luminous AGN
can be explained by a simple geometrical picture involving an accretion
disc and associated outflow. Here, we test this paradigm by 
utilising our state-of-the-art radiative transfer code to produce synthetic spectra
from simple biconical disc wind models. In particular, we expand on our previous 
work in which a benchmark model for BAL quasars was produced. 
We have conducted a limited parameter search with the aim of unifying quasar phenomenology,
and arrive at an improved model. The grid is now publicly available.
We find that a simple treatment of clumping (`microclumping') 
allows for a more realistic X-ray luminosity in the model by lowering the 
ionization parameter. We examine the X-ray properties of this new model
and find good agreement with existing X-ray samples of AGN and QSOs.
We find that the dense, X-ray heated wind 
produces strong H recombination and collisionally excited resonance 
line emission to emerge at the low inclination angles, 
which represent quasars within this unification scenario.
% We also treat Hydrogen using a `macro-atom' approach in order to 
% examine the effect of recombination on Hydrogen emission lines, and
% this results in significant line emission in Lyman~$\alpha$ and the Balmer series.
However, we are unable to reproduce the remarkably similar of line-to-continuum
ratios between BAL and non-BAL quasar composites.
This is due to a fundamental constraint arising
from the anisotropy of emission from a classical thin disc. 
We briefly explore the effect of relativistic beaming, gravitational redshift and 
light bending on the angular distribution of disc continuum emission. 
We find that these general relativistic effects do cause the disc to
emit slightly more isotropically, but this is not yet sufficient to produce a
self-consistent model. We discuss wind reprocessing as a potential solution.
Overall, our work suggests that geometric unification
involving an accretion disc wind is a promising scenario, but our results 
pose a number of difficult challenges to such a model.
% Determining the true geometry of ADWs and uncovering the true disc spectral 
% (and angular) energy distribution are key next stpng if we are to build up a 
% holistic picture of the quasar population.
\end{abstract}

@@

Macro-atoms have their ion and level populations derived from
MC rate estimators as described by M15. Previously (LK02, H13, M15),
we adopted a modified Saha approach to calculate the ionization fractions
of simple-ions. We have now improved our code to explicitly solve the 
rate equations in between ions in non-LTE. 
Not only does this
dispense with a number of small assumptions made in the modified Saha approach, 
it is also more numerically stable, 
and in principle allows the direct addition of extra physical 
processes such as Auger ionization, which would have to be approximated 
if using the previous technique.

To avoid approximating the complex AGN SED as e.g. a dilute blackbody \citep[][M15]{ML93},
we model the SED in a cell using the technique described by H13. In this scheme,
the mean intensity in a series of bands is modeled as a normalised power law in 
frequency $\nu$
\begin{equation}
J_{\nu,j}=K_{pl}\nu^{\alpha_{pl}},
\end{equation}
for a band $j$, or an exponential 
\begin{equation}
J_{\nu,j}=K_{exp} e^{(-h\nu/k T_{exp})}.
\end{equation}
Here, $K_{exp}$, $K_{pl}$, $T_{exp}$ and $\alpha_{pl}$ are spectral fit parameters
deduced from the band-limited radiation field estimators.
The ionization rate out of ion $i$ can then be written as 
\begin{equation}
R_{i,i+1}(J)= 
\displaystyle{
n_i \left(C_{i} n_e + 
\sum_{band~j=0}^{n}~
{\int_{\nu_j}^{~\nu_j+1}{J_{\nu,j}\sigma_i(\nu)\nu^{-1}d\nu}}
\right),}
\end{equation}
where $\sigma_i$ is the photoionization cross-section and $C_{i}$
represents the collisional ionization coefficient.
The recombination rate {\em into} ion $i$ is simply given by 
\begin{equation}
R_{i+1,i}(T_e) = (\alpha^i_{RR} + \alpha^i_{DR} + \alpha^i_{CR}) n_{i+1} n_e
\end{equation}
Where each $\alpha^i$ here is the recombination rate coefficient into the ground state of ion $i$.
The subscripts denote radiative, dielectronic and 
collisional (three-body) recombination respectively.
We thus neglect recombination to and from excited states in the simple-ion calculation. 
For simple-ions, we use a dilute Boltzmann equation to calculate 
the population of level $k$ in ionic stage $j$,
\begin{equation}
\frac{n_{jk}}{n_j} = \frac{W g_k}{z_j(T_R)} \exp(-E_k/kT_R).
\end{equation}
Here $z_j$ is the partition function of ionic stage $j$,
$T_R$ is the effective radiation temperature,
$E_k$ is the energy difference between level $k$ and the ground state,
and $g_k$ is the statistical weight of level $k$.

@@

The accretion disc in our model is geometrically thin, but optically thick
and thus adopt a standard multi-temperature blackbody
using a \cite{shakurasunyaev1973} temperature profile.

The emergent SED is thus determined by the specified accretion rate ($\dot{m}$)
and central BH mass ($M_{BH}$).
All photon sources in our model are assumed opaque, meaning
that photons which strike them are destroyed.
The inner radius of the disc extends to the innermost 
stable circular orbit (ISCO) of the BH. 
We assume a Schwarzchild BH with an ISCO at $6~R_G$.
The X-ray source in our model is treated as an isotropic sphere at the ISCO,
which emits r-packets according to a power law in flux with index $\alpha_X$ such that
\begin{equation}
F_X (\nu) = K \nu^{\alpha_X}.
\end{equation}
The normalisation, $K_X$ of this power law is such that it 
produces the specified 2-10~keV luminosity, $L_X$.
In addition to the disc and X-ray source, 
the wind is able to reprocess radiation. However, new 
photon packets are not produced in the wind (as in LK02). 
Instead, this reprocessing is dealt with by enforcing strict
radiative equilibrium ({\em modulo} adiabatic cooling)
via an indivisible energy packet
constraint (see Lucy 2002, M15; also section 2.3).
% \begin{equation}

@@

To take account of clumping in our outflow we adopt a simple parameterization
used in stellar wind modelling, known as {\em microclumping} \citep{hamann1998}. 
The key assumption here is that typical clump sizes
are much smaller than the typical photon mean free path, and thus the clumps are 
both geometrically and optically thin. This approach is typically 
known as microclumping and allows one to introduce a `filling factor', $f$, which is the 
fraction of the volume of the plasma filled by clumps. We can then introduce the 
density enhancement, $D$, which is simply defined as 
\begin{equation}
D = \frac{1}{f}.
\end{equation}
We then multiply all densities in the model by $D$, and all emitting volumes
by $f$, meaning that all $\rho^2$ emissivities and opacities
(such as collisional excitation and recombination) will be enhanced, 
while all $\rho$ process (such as electron scattering and bound-free opacity)
remain unchanged for a given ionization state. 
% A factor $f$ is also
% applied to the opacities such that opacities which scale only with $\rho$ are not
% increased.

@@

\section{Results}

Here we describe the results from our model, the parameters
of which are shown in table~1.
Parameters differing from the benchmark model of
H13 are highlighted with an asterisk.
This set of parameters
was arrived at by conducting a limited grid search over a 
5-dimensional parameter space involving the variables
$r_{min}$, $\theta_{min}$, $f$, $\alpha$ and $R_V$.
The full grid, including output spectral files and plots can be found at
\url{jhmatthews.github.io/quasar-wind-grid/}.
We evaluated these models qualitavely based on the following
criteria:
\begin{itemize}
\item Does the model maintain the correct ionization state and produce strong BALs?
\item Does strong line emission emerge at low inclinations?
\item Do H recombination lines appear in the spectrum?
\item Do a certain range of angles produce LoBAL features?
\item Does the model compare favourably to quasar composite spectra?
\end{itemize}
We arrived at a model which is one of the most promising,
but also broadly representative of a family of models. 
Examination of this individual model allows us to gain insight
into fundamental geometrical and physical constraints, 
as discussed in section 5, and assess the models potential
for unification.

\begin{figure*} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/wind.png}
\caption
{
Physical properties of the outflow, shown by the coloured contours.
The solid black line marks a sphere at $1000~R_G$.
The dotted lines show the $72^\circ$ and $78^\circ$ sightlines 
to the centre of the system, and illustrate that different sightlines
intersect material of different ionization states.
}
\label{fig:wind}
\end{figure*} %fullpage

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/sightlines.png}
\caption
{
The classes of sightlines and the corresponding possible 
system types in the framework of our model.
%[FIGURE NEEDS IMPROVING]
}
\label{fig:sightline}
\end{figure}

@@

Figure~\ref{fig:wind} shows the physical properties of the wind.
The wind rises slowly from the disc at first, with clumped densities
of $n_H \sim 10^{11}~\rm{cm^{-3}}$ close to the disc plane.
The flow then accelerates over a scale length of $R_V=10^{19}~\rm{cm}$
up to a terminal velocity of around $3$ times the escape velocity 
($\sim10,000~\rm{km~s^{-1}}$). This gradual acceleration means that
the wind exhibits a stratified ionization structure, with low ionization material
in the base of the wind giving way to highly ionized plasma further out.
By clumping the wind, we are able to produce the range of ionization states observed
in wuasars and BALQSOs, while adopting a realistic $2-10$ keV X-ray luminosity
of $L_{X}=10^{45}~\rm{ergs~s^{-1}}$. Without clumping, this wind would be over-ionized 
to the extent that opacities in e.g., \civ\ would be entirely negligible (see H13).

One commonly used measure of the ionization state is the ionization parameter, $U$, given by
\begin{equation}
U = \frac{4\pi}{n_H c}\int_{13.6{\rm{eV}}}^{\infty}\frac{{J_{\nu}d\nu}}{h\nu}.
\end{equation}
\noindent where $\rm{n_H}$ is the local number density of H, and $\nu$ denotes photon 
frequency. The ionization parameter represents the ratio of the number density of 
ionizing photons to the local H density. It is however, a poor measure of the 
ionization state of the resonance species such as \civ\ as it encodes no information
about the shape of the SED. In this case the X-ray photons 
are dominant in the photoionization of the UV resonance line ions. 
This explains why a factor of 100 increase in X-ray luminosity requires
a clumping factor of 0.01, even though the value of U decreases by only a factor of $\sim10$ 
compared to H13. This is also the reason for significant Lyman edge photoabsorption
at the highest inclinations (see section 4.3).

Clumping also causes the total line luminosity to increase dramatically,
as recombination and collisional excitation are both proportional to
$n_e^2$. This line emission typically emerges on the edge of the wind
nearest the central source. The location of the line emitting regions
is dependent on the ionization state, as well as the X-rays heating the plasma.
The radii of these emitting regions is important,
and can be compared to observations. As shown in ~\ref{fig:wind},
the \civ\ line in our model is typically formed between 
$100-1000~R_G$ ($\sim10^{17}-10^{18}~\rm{cm}$).
This is in rough agreement with the reverberation mapping 
results of Kaspi (2000) for the $2.6\times10^{9} M_\odot$ quasar S5 0836+71,
and also compares favourably with microlensing measurements of the size of the
\civ\ emission line region in the BALQSO H1413+117 \citep{odowd2015}.

@@

Figure 4 shows the synthetic spectrum in the UV from our model. 
We also show a comparison
to composite quasar and BALQSO spectra in figure 5.
The UV spectrum is characterised by strong BAL 
profiles at high inclinations ($> 70^\circ$). 
This highlights the first success of our model: 
clumping means the correct ionization state 
is maintained in the presence of strong X-rays, 
allowing large resonance line opacities. 
At the highest inclinations, the 
cooler, lower ionization material at the base of the wind
starts to intersect the line of sight. This produces 
multiple absorption lines lower ionization species such as \mg,
\al\ and Fe~\textsc{ii}, as well as strong Lyman continuum absorption.
The potential links to LoBALQSOs and FeLoBALQSOs are discussed in section 2.4.

The BAL profiles are often saturated, and the location in velocity space
of the strongest absorption in the profile varies with inclination.
At lower inclinations, the strongest absorption occurs at the red edge,
whereas at high inclinations (and for the strongest BALs)
the trough has a sharp edge at the terminal velocity.
This offers one potential explanation for the wide range of BALQSO absorption
line shapes (see e.g. Trump et al. 2006; Knigge et al 2008).
In addition, the line profile shape is strongly dependent 
on the density, ionization and velocity 
profiles intersected by the line of sight. Thus, small tweaks of the velocity
law and angular distributions of streamlines can dramatically alter
the shape of the line.
% the line profile shape. Figure ?? shows selection of line profiles 
% for different kinematic parameters to demonstrate this. 
% The degeneracy between parameters, and lack of available
% physical and observational constraints goes someway to showing why individual 
% fitting of line profile shapes with kinematic models is such a difficult
% exercise.

We find that the model can produce significant line emission
at low inclinations, particular in \civ, 
and the improved treatment of recombination results in a strong \la\
line. In the context of unification, this is a promising results, 
and shows that a biconical wind can produce significant 
emission at `quasar-like' angles.
However, the red wing of the BAL profiles 
is generally stronger than seen in BALQSO spectra and composites.
This is due to a combination of strong collisionally excited emission
and continuum attenuation. To illustrate the amount of 
continuum attenuation we also show the continuum from a disc-only model.
The line-to-continuum ratios at low inclinations are also
significantly affected by disc foreshortening and limb darkening.
The angular distribution of the disc radiation is clearly
crucially important in determining the emergent line equivalent widths 
(EWs) -- this is discussed further in section~5.

To assess the ability of the model to match real 
quasar spectra, we also show {\sl Sloan Digital Sky Survey} (SDSS) quasar composites from
\cite{reichard2003}, normalised to the flux at 2000\AA\ in each panel.
The model matches the \civ\ and \mg\ line profiles fairly well, 
but the \la\ profile is a factor of $\sim2$ weaker than in the composite spectrum.

This is implies that a larger recombination line formation region is required in order 
to successfully reproduce the strong \la\ emission seen in 
all quasars.

We do not produce the semi-forbidden 
intercombination lines seen in quasar spectra because we
currently do not have a treatment for semi-forbidden lines.
This is especially noticable with the strong C~\textsc{iii}]~1909\AA\, 
line in the quasar composite spectra.
The critical density of the C~\textsc{iii}]~1909\AA\, line 
is $n_e\sim10^{9.5}~\rm{cm^{-3}}$, which is higher than much of the 
outer portion of our wind. We therefore expect intercombination lines
to become important coolants in the outer portion of the wind.

@@

\begin{figure*} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/alpha_ox_both.png}
\caption
{
X-ray properties of the H13 and clumped model (text filled 
squares), plotted against monochromatic luminosity 
at 2500\AA. Also plotted are the samples considered by
Saez et al. 2012 on a similar plot; The COMBI-7 AGN and
the BQS samples Steffen et al. (2006) and the Saez et al. (2012) sample of BALQSOs.
The dotted lines show the best fit relations for non-BALQSOs from Steffen et al. (2006).
}
\label{fig:xray}
\end{figure*} %fullpage

@@

\subsection{The BALQSO fraction and wind covering factor}
\label{sec:balfrac}

>>>>

\maketitle
\begin{abstract} 
% Broad absorption lines (BALs) in the ultraviolet 
% are seen in $\sim20\%$ of quasi-stellar objects (QSOs). 
% Blue-shifted broad absorption lines (BALs) are the most direct evidence of 
% accretion disc `winds' in such systems; mass loaded outflows
% emanating from the disc that may be driven by line forces or
% magnetic processes. 
% Various unification schemes for
% quasars and luminous active galactic nuclei (AGN) have proposed
% that the broad emission line region is roughly cospatial
% with broad absorption line (BAL) gas and much of the phenomenology of luminous AGN
% can be explained by a simple geometrical picture involving an accretion
% disc and associated outflow. Here, we test this paradigm by 
% utilising our state-of-the-art radiative transfer code to produce synthetic spectra
% from simple biconical disc wind models. 
% In particular, we expand on our previous 
% work in which a benchmark model for BAL quasars was produced. 
% We have conducted a limited parameter search with the aim of unifying quasar phenomenology,
% and arrive at an improved model. The grid is now publicly available.
% We find that a simple treatment of clumping (`microclumping') 
% allows for a more realistic X-ray luminosity in the model by lowering the 
% ionization parameter. We examine the X-ray properties of this new model
% and find good agreement with existing X-ray samples of AGN and QSOs.
% We find that the dense, X-ray heated wind 
% produces strong H recombination and collisionally excited resonance 
% line emission to emerge at the low inclination angles, 
% which represent quasars within this unification scenario.
% % We also treat Hydrogen using a `macro-atom' approach in order to 
% % examine the effect of recombination on Hydrogen emission lines, and
% % this results in significant line emission in Lyman~$\alpha$ and the Balmer series.
% However, we are unable to reproduce the remarkably similar of line-to-continuum
% ratios between BAL and non-BAL quasar composites.
% This is due to a fundamental constraint arising
% from the anisotropy of emission from a classical thin disc. 
% We briefly explore the effect of relativistic beaming, gravitational redshift and 
% light bending on the angular distribution of disc continuum emission. 
% We find that these general relativistic effects do cause the disc to
% emit slightly more isotropically, but this is not yet sufficient to produce a
% self-consistent model. We discuss wind reprocessing as a potential solution.
% Overall, our work suggests that geometric unification
% involving an accretion disc wind is a promising scenario, but our results 
% pose a number of difficult challenges to such a model.
% Determining the true geometry of ADWs and uncovering the true disc spectral 
% (and angular) energy distribution are key next stpng if we are to build up a 
% holistic picture of the quasar population.
Various unification schemes for quasars and luminous active galactic nuclei (AGN) 
propose that much of their complex phenomenology can be explained by a simple 
geometrical picture involving an accretion disc and associated outflow.
Here, we test this paradigm by utilising our state-of-the-art radiative transfer 
and photoionization code to produce synthetic spectra from simple biconical disc wind models. 
In particular, we expand on our previous work by presenting a model that
exhibits many of the spectral features expected from a quasar unification model.
We find that a simple treatment of clumping (`microclumping') 
allows for a more realistic X-ray luminosity in the model,
while maintaining the ionization state necessary for strong BAL features. 
We examine the X-ray properties of this new model
and find good agreement with existing X-ray samples of AGN and quasars.
We find that the dense, X-ray heated wind 
produces strong recombination and collisionally excited line emission in, 
e.g., \civ\ and \la, to emerge at the low inclination, `Type 1 quasar-like' angles. 
The highest inclination models possess prominent \mg\ and \al\ BALs, the absorption 
features seen in LoBAL quasars. Despite these successes, we are unable to reproduce 
the remarkably uniform emission line properties 
% remarkably similar of line-to-continuum ratios apparent 
seen in BAL and non-BAL quasar composites.
This is due to a fundamental constraint arising
from the anisotropy of emission from a classical thin disc. 
We briefly explore the impact of general relativistic effects
on the angular distribution of disc continuum emission, and find that 
they are not sufficient to produce a self-consistent model.
% relativistic beaming, gravitational redshift and 
% light bending on the angular distribution of disc continuum emission. 
% We find that these general relativistic effects are not yet sufficient to produce a
% self-consistent model. 
Overall, our work suggests that geometric unification
involving an accretion disc wind is a promising scenario, but our results 
pose a number of difficult challenges to such a model.
\end{abstract}

@@

Macro-atoms have their ion and level populations derived from
MC rate estimators as described by M15. Previously (LK02, H13, M15),
we adopted a modified Saha approach to calculate the ionization fractions
of simple-ions. We have now improved our code to explicitly solve the 
rate equations between ions in non-LTE. 
This dispenses with a number of small assumptions made in the modified Saha approach, 
is more numerically stable, 
and, in principle, allows the direct addition of extra physical 
processes such as Auger ionization.

%To avoid approximating the complex AGN SED as e.g. a dilute blackbody \citep[][M15]{ML93},
We model the SED in a cell using the technique described by H13. In this scheme,
the mean intensity in a series of $n$ bands is modeled as a normalised power law in 
frequency $\nu$
\begin{equation}
J_{\nu,i}=K_{pl}\nu^{\alpha_{pl}},
\end{equation}
for a band $i$, or an exponential 
\begin{equation}
J_{\nu,i}=K_{exp} e^{(-h\nu/k T_{exp})}.
\end{equation}
Here, $K_{exp}$, $K_{pl}$, $T_{exp}$ and $\alpha_{pl}$ are spectral fit parameters
deduced from the band-limited radiation field estimators.
The ionization rate out of ion $j$ can then be written as 
\begin{equation}
R_{j,j+1}(J)= 
\displaystyle{
n_i \left(C_{j} n_e + 
\sum_{band~i=0}^{n}~
{\int_{\nu_i}^{~\nu_i+1}{J_{\nu,i}\sigma_j(\nu)\nu^{-1}d\nu}}
\right),}
\end{equation}
where $\sigma_j$ is the photoionization cross-section and $C_{j}$
represents the collisional ionization coefficient.
The recombination rate {\em into} ion $j$ is given by 
\begin{equation}
R_{j+1,j}(T_e) = (\alpha^j_{RR} + \alpha^j_{DR} + \alpha^j_{CR}) n_{j+1} n_e,
\end{equation}
where each $\alpha^j$ here is the recombination rate coefficient into the ground state of ion $j$.
The subscripts denote radiative, dielectronic and 
collisional (three-body) recombination, respectively.
We neglect recombination to and from excited states in the simple-ion calculation. 
For simple-ions, we use a dilute Boltzmann equation to calculate 
the population of level $k$ in ionic stage $j$,
\begin{equation}
\frac{n_{jk}}{n_j} = \frac{W g_k}{z_j(T_R)} \exp(-E_k/kT_R).
\end{equation}
Here $z_j$ is the partition function of ionic stage $j$,
$T_R$ is the effective radiation temperature,
$E_k$ is the energy difference between level $k$ and the ground state,
and $g_k$ is the statistical weight of level $k$.

@@

The accretion disc in our model is geometrically thin, but optically thick.
We thus treat the disc as an ensemble of blackbodies with a 
\cite{shakurasunyaev1973} effective temperature profile. 
The emergent SED is then determined by the specified accretion rate ($\dot{m}$)
and central BH mass ($M_{BH}$).
All photon sources in our model are assumed to beopaque, meaning
that photons which strike them are destroyed.
The inner radius of the disc extends to the innermost 
stable circular orbit (ISCO) of the BH. 
We assume a Schwarzchild BH with an ISCO at $6~r_G$, where 
$r_G = GM_{BH}/c^2$ is the gravitational radius.
For a $10^9~M_\odot$ black hole, this is equal to $8.8\times10^{14}~{\rm cm}$ 
or $\sim10^{-4}~{\rm pc}$.

The X-ray source in our model is treated as an isotropic sphere at the ISCO,
which emits r-packets according to a power law in flux with index $\alpha_X$, of the form
\begin{equation}
F_X (\nu) = K \nu^{\alpha_X}.
\end{equation}
The normalisation, $K_X$ of this power law is such that it 
produces the specified 2-10~keV luminosity, $L_X$.
In addition to the disc and X-ray source, 
the wind is able to reprocess radiation. However, new 
photon packets are not produced in the wind (as in LK02). 
Instead, this reprocessing is dealt with by enforcing strict
radiative equilibrium ({\em modulo} adiabatic cooling; see section~2.3)
via an indivisible energy packet
constraint (see Lucy 2002, M15).
% \begin{equation}

@@

To allow for clumping in our outflow we adopt a simple approximation
used extensively in stellar wind modelling, known as 
{\em microclumping} \citep{hamann1998}. The key assumption here is that 
typical clump sizes are much smaller than the typical photon mean free path, 
and thus the clumps are 
both geometrically and optically thin. This approach 
allows one to introduce a `filling factor', $f$, which is the 
fraction of the volume of the plasma filled by clumps. The intra-clump
medium is assumed to be a vacuum. We can then introduce the 
density enhancement, $D$, which is defined as 
\begin{equation}
D = \frac{1}{f}.
\end{equation}
We then multiply all densities in the model by $D$, and all emitting volumes
by $f$. This has the effect of enhancing all emissivities and opacities
that scale with the square of density (such as collisional excitation and recombination) 
by a factor $D$. All processes that scale linearly with density 
(such as electron scattering and bound-free opacity)
will remain unchanged for a given ionization state. 
% A factor $f$ is also
% applied to the opacities such that opacities which scale only with $\rho$ are not
% increased.

@@

\section{Results and Discussion}

Here we describe the results from our model, the parameters
of which are shown in table~1.
Parameters differing from the benchmark model of
H13 are highlighted with an asterisk.
This set of parameters
was arrived at by conducting a limited grid search over a 
5-dimensional parameter space involving the variables
$r_{min}$, $\theta_{min}$, $f$, $\alpha$ and $R_V$.
The full grid, including output spectral files and plots can be found at
\url{jhmatthews.github.io/quasar-wind-grid/}.
We evaluated these models qualitatively based on the following
criteria:
\begin{itemize}
\item Does the model maintain the right ionization state to produce strong BALs?
\item Does significant line emission emerge at low inclinations?
\item Do H recombination lines appear in the spectrum?
\item Do a certain range of angles produce LoBAL features?
\item Does the model compare favourably to quasar composite spectra?
\end{itemize}
In this section, we present one of the most promising models and discuss
the various successes and failures.
This allows us to gain insight into fundamental geometrical 
and physical constraints and assess the potential for unification.

\begin{figure*} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/link.png}
\caption
{
Physical properties of the outflow, shown by the coloured contours.
The solid black line marks a sphere at $1000~r_G$.
The dotted lines show the $72^\circ$ and $78^\circ$ sightlines 
to the centre of the system, and illustrate that different sightlines
intersect material of different ionization states.
}
\label{fig:wind}
\end{figure*} %fullpage

@@

Figure~\ref{fig:wind} shows the physical properties of the wind.
The wind rises slowly from the disc at first, with clumped densities
of $n_H \sim 10^{11}~\rm{cm^{-3}}$ close to the disc plane.
The flow then accelerates over a scale length of $R_V=10^{19}~\rm{cm}$
up to a terminal velocity equal to the escape velocity at the streamline base
($\sim10,000~\rm{km~s^{-1}}$). This gradual acceleration means that
the wind exhibits a stratified ionization structure, with low ionization material
in the base of the wind giving way to highly ionized plasma further out.
By clumping the wind, we are able to produce the range of ionization states observed
in quasars and BALQSOs, while adopting a realistic $2-10$ keV X-ray luminosity
of $L_{X}=10^{45}~\rm{ergs~s^{-1}}$. Without clumping, this wind would be over-ionized 
to the extent that opacities in e.g., \civ\ would be entirely negligible (see H13).

One common way to quantify the ionization state of a plasma
is through the ionization parameter, $U_H$, given by
\begin{equation}
U_H = \frac{4\pi}{n_H c}\int_{13.6{\rm{eV}}}^{\infty}\frac{{J_{\nu}d\nu}}{h\nu}.
\end{equation}
\noindent where $\rm{n_H}$ is the local number density of H, and $\nu$ denotes photon 
frequency. Shown in figure~\ref{fig:wind},
the ionization parameter is a useful measure of the global ionization state,
as it represents the ratio of the number density of 
H ionizing photons to the local H density.
It is, however, a poor representation of the 
ionization state of species such as \civ\ as it encodes no information
about the shape of the SED. In our case, the X-ray photons 
are dominant in the photoionization of the UV resonance line ions. 
This explains why a factor of 100 increase in X-ray luminosity requires
a clumping factor of 0.01, even though the value of $U_H$ decreases by only a factor of $\sim10$ 
compared to H13. This is also the reason for significant Lyman edge photoabsorption
at the highest inclinations (see section 4.3).

Clumping also causes the total line luminosity to increase dramatically,
as recombination and collisional excitation are both proportional to
$n_e^2$. This line emission typically emerges on the edge of the wind
nearest the central source. The location of the line emitting regions
is dependent on the ionization state, as well as the X-rays heating the plasma.
The radii of these emitting regions is important,
and can be compared to observations. As shown in figure~\ref{fig:wind},
the \civ\ line in our model is typically formed between 
$100-1000~r_G$ ($\sim10^{17}-10^{18}~\rm{cm}$).
This is in rough agreement with the reverberation mapping 
results of Kaspi (2000) for the $2.6\times10^{9} M_\odot$ quasar S5 0836+71,
and also compares favourably with microlensing measurements of the size of the
\civ\ emission line region in the BALQSO H1413+117 \citep{odowd2015}.

@@

Figure~\ref{fig:uvspec} shows the synthetic spectrum in the UV from our model. 
We also show a comparison to composite quasar and BALQSO spectra.
We show a cartoon illustrating how geometric effects determine
the output spectra in figure~\ref{fig:sightline}.

\subsubsection{Broad absorption lines}

The UV spectrum is characterised by strong BAL 
profiles at high inclinations ($> 70^\circ$). 
This highlights the first success of our model: 
clumping means the correct ionization state 
is maintained in the presence of strong X-rays, 
allowing large resonance line opacities. 
At the highest inclinations, the 
cooler, lower ionization material at the base of the wind
starts to intersect the line of sight. This produces 
multiple absorption lines in lower ionization species such as \mg,
\al\ and Fe~\textsc{ii}. The potential links to LoBALQSOs and 
FeLoBALQSOs are discussed in section 2.4.

The high ionization BAL profiles are often saturated, and the location in velocity space
of the strongest absorption in the profile varies with inclination.
At lower inclinations, the strongest absorption occurs at the red edge,
whereas at high inclinations (and for the strongest BALs)
the trough has a sharp edge at the terminal velocity.
This offers one potential explanation for the wide range of BALQSO absorption
line shapes (see e.g. Trump et al. 2006; Knigge et al 2008).
In addition, the line profile shape is strongly dependent 
on the density, ionization and velocity 
profiles intersected by the line of sight. Thus, small tweaks of the velocity
law and angular distributions of streamlines can dramatically alter
the shape of the line.

Nonblack saturation is observed in the absorption troughs of BALQSOs \citep{arav1999a,arav1999b}.
This can be caused either by partial covering of the continuum
source or by scattered contributions to the BAL troughs, necessarily
from an opacity source not cospatial with the BAL forming region.
The scattered light explanation is supported by spectropolarimetry results
\citep{lamy2000}. Our spectra do not show nonblack saturation.
Instead, we find black, saturated troughs at angles $i > 73^\circ$, and the BALs
are non-saturated at lower inclinations. The reasons for this are readily apparent. 
First, the microclumping assumption does not allow for 
porosity in the wind, meaning that it does not naturally produce
a partial covering absorber. To do this, an alternative approach
such as {\em macroclumping} would be required \citep[e.g.][]{surlan2012,hamann2008}.
Second, our wind does not have a significant scattering 
contribution along sightlines which do not pass through the BAL region,
meaning that any scattered contribution is obscured by the saturated troughs.
This suggests that either the scattering cross-section of the wind must
be increased (with higher mass loss rates or covering factors), or 
that an additional source of electron opacity is required, potentially
in a polar direction above the disc.

\subsubsection{Broad emission lines}

We find that the model can produce significant line emission
at low inclinations, particular in \civ, 
and the improved treatment of recombination results in a strong \la\
line. In the context of unification, this is a promising result, 
and shows that a biconical wind can produce significant 
emission at `quasar-like' angles. To assess the ability of the model to match real 
quasar spectra, we also show {\sl Sloan Digital Sky Survey} (SDSS) quasar composites from
\cite{reichard2003}, normalised to the flux at 2000\AA\ in each panel.
We do not produce the semi-forbidden 
intercombination lines seen in quasar spectra because we
currently do not have a treatment for semi-forbidden lines.
This is especially noticable with the strong C~\textsc{iii}]~1909\AA\, 
line in the quasar composite spectra.
The critical density of the C~\textsc{iii}]~1909\AA\, line 
is $n_e\sim10^{9.5}~\rm{cm^{-3}}$, which is higher than much of the 
outer portion of our wind. We therefore expect intercombination lines
to become important coolants in the outer portion of the wind.

The model produces strong emission lines in \civ, \nv\ and \la,
as well as a weak \mg\ line. The shapes and widths of these lines
match the composites fairly well. However, the line-to-continuum ratios 
at low inclinations in our model are significantly weaker than the quasar 
composites. Increasing the density of the outflow, by altering the mass 
loss rate  or velocity law, can produce more line emission.
However, the red wing of the BAL profiles is generally stronger than 
seen in BALQSO spectra and composites. This illustrates a fundamental 
problem with a geometric unification model such as this:
that the line-to-continuum ratios at low inclinations are
significantly affected by disc foreshortening and limb darkening.
The angular distribution of the disc radiation is clearly
crucially important in determining the emergent line ratios.

\subsubsection{The Disc SED}
\label{discsed}

When comparing BALQSO and quasar composites, it is apparent
that they possess remarkably similar line strengths and widths 
\citep[e.g.][]{reichard2003}.
This presents a challenge to our model, as well as the geometric 
unification picture in general.
Limb darkening and foreshortening causes the disc continuum emission to be 
strongly anisotropic, yet the line emission in our models is much more isotropic. 
This has the effect of enhancing the line-to-continuum ratios at high inclinations. 
To construct a scenario where emission 
line equivalent widths are comparable at all inclinations requires 
significant fine-tuning when considering a classical thin disc. 
An alternative, simpler solution is that the emergent continuum is roughly isotropic.

General relativistic effects -- specifically, light bending
and relativistic beaming -- can cause 
the accretion disc SED to become more isotropic \citep[e.g.][]{zhang1997,munozdarias2013}.
To generate GR disc spectra, we use the code \agn\ \citep{hubeny2000,davishubeny2006,davis2007}. 
The output flux at three different wavelengths
as a function of inclination for an \agn\ model with the same disc and BH parameters
as our clumpy wind model is shown in figure~\ref{fig:f2000}.
The effects of GR on an AGN disc are much less extreme 
in the UV portion of the spectrum than the calculations
by \cite{zhang1997} for the X-rays in X-ray binaries.
As a result, the disc emission is still strongly anisotropic.
GR alone therefore cannot explain the line ratio trends in quasars by making the disc
emit more isotropically.
This supports the findings of \cite{risaliti2011}, who find that EW distributions 
in quasars are consistent with anisotropic emission from optically thick, 
disc-like structures for {\em both} the continuum source and BLR. 
We will explore these ideas further in a future study.

@@

\begin{figure*} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/lx.png}
\caption
{
X-ray properties of the H13 (purple squares) and clumped model (red
squares), plotted against monochromatic luminosity 
at 2500\AA. The points are labeled according to inclination; angles
$>70^\circ$ correspond to BALs in our scheme.
Also plotted are the samples considered by Saez et al. 2012 on a similar plot; 
The COMBI-7 AGN and the BQS samples Steffen et al. (2006) and the Saez et al. (2012) 
sample of BALQSOs. The dotted lines show the best fit relations for non-BALQSOs 
from Steffen et al. (2006).
}
\label{fig:xray}
\end{figure*} %fullpage

@@

% \subsection{The BALQSO fraction and wind covering factor}
% \label{sec:balfrac}
